Synthesizing Symmetric Lenses

### 2018-07-01 papers/SymmetricLenses/main.tex
Finished pass over section 2

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
      \hline
      \texttt{company.mgmt} data file & \texttt{company.hr} data file \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{vlightyellow}\texttt{company.mgmt} regular expression & 
      \cellcolor{vlightgreen}\texttt{company.hr}  regular expression\\
      \hline
      \cellcolor{vlightyellow}\listingC & \cellcolor{vlightgreen}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Example data files and corresponding regular expressions
  used by management and HR, respectively, to represent employee salaries and health
  insurance companies.  }  
  \label{fig:minimized-representations}
\end{figure}

In this section, we give an overview of simple symmetric lenses and
our corresponding synthesis algorithm using an example drawn from a
hypothetical U.S. company.  In this company, management and
human resources (HR) store information about employees in separate files:
management records the names of employees and their salaries while
HR stores the names of employees and their health
insurance providers.  Figure~\ref{fig:minimized-representations} gives
examples of the two types of files and regular expressions describing
the two formats.
The company uses simple symmetric lenses to keep these files synchronized.
If management adds a new employee, ``Chris Roe: 32500'',  a lens adds 
the corresponding entry ``Chris Roe, UNK'' to HR's file.  The
sentinel value ``UNK'' represents the fact that the employee's
insurance company is currently unknown.  A similar update happens if
HR adds a new employee before management, in which case the sentinel
value ``N/A'' reflects the unknown salary information. 
Furthermore, if HR
corrects an error in an employee's name, say changing ``John Public''
to ``Jon Public'', the lens mirrors that name change in management's
file. Not all the data is mirrored, however. The management file 
is not updated in response to insurance changes, and the HR file
is oblivious to salary changes.  Simple symmetric lenses are
appropriate for keeping these two files synchronized because they have
a mix of shared and unshared information.

@@

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by types $\Regex$
and $\RegexAlt$ 
is defined by four functions: 
\begin{enumerate}
\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
\end{enumerate}
\noindent
that collectively satisfy four ``round-tripping'' laws: 
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{s})}{s} = s\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{t})}{t} = t\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{s}{t})}{s} = s\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{t}{s})}{t} = t
\end{gather}
\noindent
The two \emph{create} functions provide default values for times when
data appears on one side but not the other (eg, the ``N/A'' salary
entry added to the management file when HR inserts a new employee).
The two \emph{put} functions propagate edits from one format to the
other by combining the change from one file with the previous value
from the other file. The roundtripping laws guarantee that identify edits to one file do
not change the data in the other file.

@@

\subsection{Simple Symmetric Lens Combinators} 
\ksf{This section needs more polishing}
Generally, we do not expect programmers to define simple symmetric lenses by
writing the four functions by hand and showing they satisfy the
roundtripping laws.  Instead, we provide simple symmetric lens
combinators to build up the requisite lenses from simpler ones.
We explain these combinators by using them to define a lens between
the two employee data formats in Figure~\ref{fig:minimized-representations}. 
The finished lens, which appears in Figure~\ref{fig:example_lens}, has
the type 
\lstinline{emp_salaries} $\Leftrightarrow$ \lstinline{emp_insurance}, where
\lstinline{emp_salaries} and \lstinline{emp_insurance} are regular
expressions defined in Figure~\ref{fig:minimized-representations}.

The simplest combinator is the identity lens, which takes as an argument a
regular expression, and performs the 
identity function on strings in that regular expression. The identity lens fully
matches the data formats: when one data format is updated, the other format gets
the exact same value. In our example, when a name is updated, it is also
expected to be updated in the other format. The identity lens propagates that
information.
%
\begin{lstlisting}
id(name) : name $\Leftrightarrow$ name
\end{lstlisting}
%

In contrast to the identity lens, the disconnect lens leaves data
unaligned. The disconnect lens takes four arguments: two regular 
expressions and two strings. The regular expressions specify the
formats of the data on the two sides while the strings provide default
values.  In our extended example, salaries and
insurance companies unrelated, so we use a disconnect lens.  The
defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disc(salary,company,"N/A","UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
When building a disconnect lens between two strings, one of which is merely the
empty string, the lenses \lstinline{ins} and \lstinline{del} can instead be
used, where \lstinline{ins($t$) = disc("",$t$,"",$t$)} and
\lstinline{del($s$) = disc($s$,"",$s$,"")}.

Finally, there are a number of lenses that compositionally build up new lenses
from old ones. 
For example, data can be concatenated, and so we provide
the \lstinline{concat} combinator that sequences lenses, usually
written using an infix notation \lstinline{$\Lens_1$ . $\Lens_2$}:
%
\begin{lstlisting}
id(name) . ins(",") . id(name) . del(": ") . ins(",")
\end{lstlisting}
% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%

Data can also appear in multiple forms. For example,
a list of employees with insurance information can either be empty or
be a sequence of employee/insurance records: \ksf{This example is
vacuous as you could eliminate the or in favor of the *.  Do somethign
else to introduce or?}
%
\begin{lstlisting}
or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
\end{lstlisting}

Lastly, the \IterateLens lens synchronizes repeated data, aligning the number of
times the data is repeated. The \IterateLens combinator takes a lens as an
argument and synchronizes each element of the iteration with that lens. For
example, if \lstinline{employee_lens} synchronizes a single employee's data,
iterating that lens builds a lens that synchronizes a list of employees.
%
\begin{lstlisting}
iterate(id("\n") . employee_lens) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disc(salary,company,"N/A","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
let emp_lens = or(id(""),employees_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

Using these combinators, we can build up a lens that synchronizes
management and HR employee files.

\ksf{This paragraph needs to be updated, if nothing else, to reflect
the new example. Also, we haven't mentioned any typing rules so the
first sentence doesn't make sense.  Perhaps the whole topic could be
defered to later where we can explain why such a rule is necessary?}
This lens uses one typing rule not previously
mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
a regular expression with an outermost disjunction, it is not syntactically
equivalent to one. Without type equivalence, the provided lens wouldn't actually
be well-typed, but type equivalence states that if a lens is typed between a
pair of regular expressions, it is also typed between an equivalent pair.

Note that, while the synchronization task is quite simple, the
requisite lens is relatively complex. Furthermore, this complexity is
inherent in manually writing simple symmetric lenses because the
lenses must precisely describe the portions of the data that should be
connected and those that should be disconnected.  For disconnected
pieces, the lens must further specifiy appropriate defaults.

@@

To alleviate these difficulties, we wish to synthesize symmetric lenses---given
a pair of regular expressions and a set of input-output examples, we want to
find a symmetric lens expression that is typed between the regular expression
pair, and satisfies the input/output examples. We call such a lens
a \emph{satisfying lens}.
In this example, we wish to
synthesize a lens from \lstinline{emp_salaries} and \lstinline{emp_insurance},
using the example data in Figure~\ref{fig:minimized-representations} as an
input-output example from which to draw the required defaut values for 
\Disconnect lenses.

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disc(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the synchronized files from
Figure~\ref{fig:minimized-representations} as an input-output example helps
little. The default strings given to \Disconnect merely need to be the strings
in the input-output examples, and the lens will still satisfy the specification.
While the disconnect lens is prohibited by two examples, or by a carefully
chosen input-output example on \PutR or \PutL, this problem continues presenting
itself for the sublenses, and is highlighted even more on larger
formats. 
\ksf{I don't understand the above.  We shouldn't explain the issue in
terms of \PutR and \PutL as we are focused on the combinators rather
than the raw functions.}
While
we could ask programmers to provide detailed and precise specifications for their
synthesis tasks, we would prefer the algorithm to choose the right lens
automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
synthesize the \emph{best} satisfying lens.

But what is the best satisfying lens? We propose that the
best satisfying lens is the satisyfing lens that connects as much data
as possible between the two formats. More formally, we
propose that the best satisfying lens is the satisfying lens that 
minimizes the expected number of bits required to recover data in
one format from synchronized data in the other.  
In more detail, we say that two strings $s$ and $t$
are \emph{synchronized according to lens $l$} if 
$l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$ where the notation $l.\PutR$
extracts the \PutR function from simple symmetric lens $l$; similarly
for \PutL.
We say that we can \emph{recover} $s$ from $t$ using bits $b$ and lens
$l$ if we can reconstruct $s$ from $t$, $b$, and $l$.  For example,
given the \IdentityLens{} lens, we can recover $s$ from $t$ using
no bits because in this case, $s$ is just $t$.  In contrast, given
the \Disconnect{} lens, we need enough bits to fully encode $s$ in
order to recover it from $t$ because $t$ contains no useful
information.  This calculation shows that if both \IdentityLens{}
and \Disconnect{} are satisfying lenses, then \IdentityLens{} is
better.

The expected number of bits required to recover some data is a
well-known concept in information theory: it is the \emph{entropy} of
the data. Calculating the 
entropy of some data requires a probability distribution over the
space of possible values the data can take on.  Specifically,
given a set $S$ and a probability distribution over that set $P
\OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
  S}P(s)\Log_2P(s)$.

@@

This infrastructure gives us what we need to find the best
satisfying lens from a collection of satisfying lenses.  All such
lenses must synchronize data between the same two 
stochastic regular expressions $\Regex$ and $\RegexAlt$.
For any such lens $\Lens$, we can calculate the expected number of
bits required 
to recover a string $s$ in $\LanguageOf{\Regex}$ from a synchronized
string $t$ in $\LanguageOf{\RegexAlt}$. 
This expected number of bits is just the conditional entropy of
$\Regex$ given $\RegexAlt$ and $\Lens$. 
To ensure symmetry, the entropy we assign to $\Lens$ is the sum of 
the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ with 
the conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$.

@@

\ksf{Should we be talking about ``best'' or just ``good'' or ``low''?
There are typically a collection of ``best'' lenses, and we choose the
one that requires the simplest permutation.}
Now that we have a notion of best, we need a way to search for that
best lens.  Previous work on synthesis~\ksf{add citations} has shown
that we can use types to dramatically restrict the search space and
improve the effectiveness of synthesis.  In our setting, types are
pairs of stochastic regular expressions, which come with a rich
algebraic structure that makes each regular expression semantically
equal to infinitely many others.  To tame that search space and
leverage the benefits of type-directed synthesis, we split our
algorithm into two communicating synthesizers following earlier
work~\cite{quotient-lens-synthesis}.  The first
synthesizer, \RXSearch, navigates the space of semantically equivalent
regular expressions by iteratively applying rewrite rules that
preserve both the semantics and the probability
distributions of those stochastic regular expressions. \RXSearch ranks
pairs of stochastic regular expressions 
by the number of rewrite rule applications required to obtain the pair
from the stochastic regular expressions given as input. It passes the
pairs off to the second synthesizer, \GreedySynth, in rank order.

At a high level, \GreedySynth looks for lenses between a given pair of
stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the
best conditional entropy by performing a type-directed search. 
In more detail, \GreedySynth first uses type equalities based on semiring 
equivalences (ignoring star semiring equivalences) to normalize the
stochastic regular expressions provided by \RXSearch into 
\emph{stochastic n-ary DNF regular expressions}.  
Then, \GreedySynth uses the syntax of these n-ary DNF regular
expressions to find highly normalized lenses, in a
form we call \emph{simple symmetric n-ary DNF lenses}.  These lenses include
neither a composition operator nor a type equivalence rule.  These
restrictions mean that each simple symmetric n-ary DNF lens is typed
by a single pair of stochastic n-ary DNF regular expressions and they
make \GreedySynth's search space finite.  The variations that \GreedySynth
considers include choosing between \IdentityLens{} and  \Disconnect
lenses and deciding how to match multiple occurrences of the same kind
of data, \EG, the two occurrences of \lstinline{name} in the
management and HR data formats (we want to map first names to first
names and last names to last names). 
Finally, \GreedySynth converts this lens 
back into a simple symmetric lens by converting the n-ary forms into
the binary forms provided in the surface language.

As an example of the partnership between \RXSearch and \GreedySynth,
consider the lens shown in Figure~\ref{fig:example_lens}. 
This lens is \emph{not} in the set of lenses 
characterized by \lstinline{emp_salaries} and \lstinline{emp_insurance}, as that
typing derivation requires unrolling \lstinline{("\n" . emp_salary)$^*$} within
\lstinline{emp_salaries} into \lstinline{"" | (("\n" . emp_salary) . ("\n" . emp_salary)$^*$)}.

Our strategy involving communicating synthesizers permits both a way of
enumerating classes and efficiently searching through them, but it poses a
problem: when do we stop? While we may have found a great lens between two regular
expressions, a different regular expression pair may provide an even better
lens. Moreover, we may find a great lens characterized by
\lstinline{emp_salaries} and \lstinline{emp_insurances} that is not the
desired lens shown in Figure~\ref{fig:example_lens}.  The
tension is that we want to find a good lens, but we also want to give quick feedback to
the user. After finding the lens in Figure~\ref{fig:example_lens}, we dont want
the algorithm to spend much longer searching for better lenses.

\saz{The following is hard to follow} Eventually, we cut the proposal-synthesis
loop short. Before proposing new types \RXSearch evaluates whether it should
continue proposing regular expression pairs, or whether it is content with lens
of lowest cost seen thus far. If there are a large number of regular expression
pairs that are equally dissimilar to the originals or if the current lens is
very well synchronized, \RXSearch stops. By doing this \RXSearch continues
searching for better lenses, but does not tackle huge search spaces unless it
needs to. In this way, the algorithm can provide lenses quickly to the users,
which users can refine with additional examples.

\afm{calculations for cost of various lenses between \lstinline{emp_salary} and
  \lstinline{emp_row}} Consider the lens \lstinline{single_emp_map} between
\lstinline{emp_salary} and \lstinline{emp_row}. Given this lens, the expected
entropy of recovering the left format from the right
($\EntropyOf{\lstinline{emp_salary} \Given \lstinline{emp_row},
  \lstinline{single_emp_map} }$) is the sum of the expected entropy of
recovering each of its subcomponents, $\EntropyOf{\lstinline{emp_salary} \Given
  \lstinline{emp_row}, \lstinline{single_emp_map} } = \EntropyOf{\lstinline{name
    . " " . name} \Given \lstinline{name . "," . name}, \lstinline{name_map}
}$\\ + $\EntropyOf{\lstinline{": " . salary} \Given
  \lstinline{"," . company}, \lstinline{disc(salary,company,"UNSET","UNSET")}
}$.

>>>>

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
      \hline
      \texttt{company.mgmt} data file & \texttt{company.hr} data file \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{vlightyellow}\texttt{company.mgmt} regular expression & 
      \cellcolor{vlightgreen}\texttt{company.hr}  regular expression\\
      \hline
      \cellcolor{vlightyellow}\listingC & \cellcolor{vlightgreen}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Hypothetical example data files and corresponding regular expressions
  used by management and HR at a US company to represent employee salaries and health
  insurance providers, respectively. \ksf{Why
  is \lstinline{emp_salaries} written with an or instead of just using
  the star operator? Can we simplify?} }  
  \label{fig:minimized-representations}
\end{figure}

In this section, we give an overview of simple symmetric lenses and
our corresponding synthesis algorithm using an example drawn from a
hypothetical U.S. company.  In this company, management and
human resources (HR) store information about employees in separate files:
management records the names of employees and their salaries while
HR stores the names of employees and their health
insurance providers.  Figure~\ref{fig:minimized-representations} gives
examples of the two types of files and regular expressions describing
the two formats.
The company uses simple symmetric lenses to keep these files synchronized.
If management adds a new employee, ``Chris Roe: 32500'',  a lens adds 
the corresponding entry ``Chris Roe, UNK'' to HR's file.  The
sentinel value ``UNK'' represents the fact that the employee's
insurance company is currently unknown.  A similar update happens if
HR adds a new employee before management, in which case the sentinel
value ``N/A'' reflects the unknown salary information. 
Furthermore, if HR
corrects an error in an employee's name, say changing ``John Public''
to ``Jon Public'', the lens mirrors that name change in management's
file. Not all the data is mirrored, however. The management file 
is not updated in response to insurance changes, and the HR file
is oblivious to salary changes.  Simple symmetric lenses are
appropriate for keeping these two files synchronized because the files have
a mix of shared and unshared information.

@@

\subsection{Simple Symmetric Lenses}
A simple symmetric lens between formats described by types $\Regex$
and $\RegexAlt$ 
is defined by four functions

%\begin{enumerate}
%\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
%\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
%\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
%\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
%\end{enumerate}

\begin{gather}
 \tag{1}
 \mbox{$\CreateR \OfType \Regex \rightarrow \RegexAlt$}\\
 \tag{2}
 \mbox{$\CreateL \OfType \RegexAlt \rightarrow \Regex$}\\
 \tag{3}
 \mbox{$\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$}\\
 \tag{4}
 \mbox{$\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$}
\end{gather}

\noindent
that collectively satisfy four ``round-tripping'' laws: 
\begin{gather}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{s})}{s} = s\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{t})}{t} = t\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{s}{t})}{s} = s\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{t}{s})}{t} = t
\end{gather}
\noindent
The two create functions provide default values for times when
data appears on one side but not the other (\EG, the ``N/A'' salary
entry added to the management file when HR inserts a new employee).
The two put functions propagate edits from one format to the
other by combining the change from one file with the previous value
from the other file. The roundtripping laws guarantee that identify edits to one file do
not change the data in the other file.

@@

\subsection{Simple Symmetric Lens Combinators} 
Generally, we do not expect programmers to define simple symmetric lenses by
writing the four functions by hand and showing they satisfy the
roundtripping laws.  Instead, we provide simple symmetric lens
combinators to build up the requisite lenses from simpler ones.
We briefly explain a selection of these combinators by using them to
define a lens between 
the two employee data formats in Figure~\ref{fig:minimized-representations}. 
The finished lens, which appears in Figure~\ref{fig:example_lens}, has
the type 
\lstinline{emp_salaries} $\Leftrightarrow$ \lstinline{emp_insurance}, where
\lstinline{emp_salaries} and \lstinline{emp_insurance} are regular
expressions defined in Figure~\ref{fig:minimized-representations}.

The simplest combinator is the identity lens \IdentityLens, which takes as an
argument a regular expression $\Regex$ and performs an identity mapping on data
matching $\Regex$. The identity lens fully connects the
corresponding data in the two formats: when one format is
updated, the other format gets the exact same value. In our example,
when a \lstinline{name} is updated in one file, the identify lens ensures it is
also updated in the other file.

%
\begin{lstlisting}
id(name) : name $\Leftrightarrow$ name
\end{lstlisting}
%

In contrast to the identity lens, the disconnect lens \Disconnect leaves data
unconnected. The disconnect lens takes four arguments: two regular 
expressions and two strings. The regular expressions specify the
formats of the corresponding data on the two sides while the strings provide default
values.  In our extended example, 
the \lstinline{salary} and \lstinline{company} fields are independent,
so we use the disconnect lens to relate them.
The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disc(salary, company, "N/A", "UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively. \ksf{Are these in the right order?}
\begin{lstlisting}
 ins($t$) = disc("", $t$, "", $t$)
 del($s$) = disc($s$, "", $s$, "")
\end{lstlisting}

Finally, there are a number of standard lenses that compositionally
construct more complex lenses from simpler ones.  These combinators include 
\lstinline{concat} for concatenation, 
\lstinline{or} for variation, and
\IterateLens for sequences.
We usually write \lstinline{concat} using an infix
notation \lstinline{$\Lens_1$ . $\Lens_2$}
as in:
%
\begin{lstlisting}
id(name) . ins(",") . id(name) . del(": ") . ins(",")
\end{lstlisting}
% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%
which appears in the definition of the \lstinline{name_lens} in our
running example. 
%\ksf{I don't think this example is necessary.}
%Data can also appear in multiple forms. For example,
%a list of employees with insurance information can either be empty or
%be a sequence of employee/insurance records: \ksf{This example is
%vacuous as you could eliminate the or in favor of the *.  Do somethign
%else to introduce or?}
%
%\begin{lstlisting}
%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
%\end{lstlisting}
The \IterateLens lens also appears in our running example, where it
synchronizes data for a list of employees when given a
lens \lstinline{employee_lens} that synchronizes  data for a single employee:
%
\begin{lstlisting}
iterate(id("\n") . employee_lens) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disc(salary,company,"N/A","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens, written in our language for building simple symmetric lenses
    on string data, that synchronizes management and HR employee files.
    Note that we can annotate our lenses with types, to confirm the lens
    operates on the languages of the regular expressions. In this work, we
    synthesize the lens itself from such regular expressions.}
  \label{fig:example_lens}
\end{figure}

% Using these combinators, we can build up a lens that synchronizes
% management and HR employee files.

We provide typing rules for each of the simple symmetric combinators,
most of which are syntax directed.  There is one additional
typing rule, which allows us to use equivalence relations on regular
expressions to convert the type of any lens to a semantically equal
but syntactically different pair of regular expressions.  The flexibility this
type equivalence rule provides is useful when a different
representation of a regular expression would facilitate synthesis.

%%\ksf{Do we need to use
%%type equivalence to type the employee example?  If so, perhaps we
%%should explain her}
%%
%%This lens uses one typing rule not previously
%%mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
%%a regular expression with an outermost disjunction, it is not syntactically
%%equivalent to one. Without type equivalence, the provided lens wouldn't actually
%%be well-typed, but type equivalence states that if a lens is typed between a
%%pair of regular expressions, it is also typed between an equivalent pair.

Note that, while the synchronization task for our running example is quite simple, the
requisite lens is relatively complex. Furthermore, this complexity is
inherent in manually writing simple symmetric lenses because the
lenses must precisely describe the portions of the data that should be
connected and those that should be disconnected.  For disconnected
pieces, the lens must further specifiy appropriate defaults.

@@

To alleviate the difficulty of manually writing simple symmetric
lenses, we wish to synthesize them instead. Given 
a pair of regular expressions and a set of input-output examples, we want to
find a simple symmetric lens that is typed by the regular expression
pair and satisfies the input/output examples. We call such a lens
a \emph{satisfying lens}.
In our running example, we wish to
synthesize a lens between \lstinline{emp_salaries} and \lstinline{emp_insurance},
using as an input-output example the data in
Figure~\ref{fig:minimized-representations}.  The algorithm uses such
examples to find suitable defaults for \Disconnect lenses and to
figure out how to match data that could be synchronised in multiple
ways.

A challenge is that the simple symmetric lens combinators permit many
well-typed lenses
between a given pair of regular expressions. For example, 
Figure~\ref{fig:example_lens} gives one possible lens between regular
expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
but 
%
\begin{lstlisting}
disc(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
\end{lstlisting}
%
is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
language of $\Regex$). Furthermore, adding the input-output example in
Figure~\ref{fig:minimized-representations} 
helps, but only a little, because it requires many examples to rule
out all the possible occurrences of nested \Disconnect lenses,
particularly in complex formats. 
%\ksf{I don't understand the following.  We shouldn't explain the issue in
%terms of \PutR and \PutL as we are focused on the combinators rather
%than the raw functions.}
%The default strings given to \Disconnect merely need to be the strings
%in the input-output examples, and the lens will still satisfy the
%specification.  While the disconnect lens is prohibited by two
%examples, or by a carefully chosen input-output example on \PutR
%or \PutL, this problem continues presenting itself for the sublenses,
%and is highlighted even more on larger formats.
While
we could ask programmers to provide detailed and precise specifications for their
synthesis tasks, we would prefer the algorithm to choose the right lens
automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
synthesize a satisfying lens that is likely to please the user.

But what is a likely satisfying lens? We propose that a satisfying
lens is more likely if it connects more data between the two
formats, \IE, by using an identity rather than a disconnect lens.  The
likeliest satisfying lenses would connect as much data as
possible. More formally, we propose that the likelihood of a
satisfying lens is the expected number of bits required to recover
data in one format from synchronized data in the other.  Higher
likelihoods correspond to fewer numbers of bits. In more
detail, we say that two strings $s$ and $t$ are \emph{synchronized
according to lens $l$} if $l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$.\footnote{
The notation $l.\PutR$ extracts the \PutR function from simple
symmetric lens $l$; similarly for \PutL.}  We say that we
can \emph{recover} $s$ from $t$ using bits $b$ and lens $l$ if we can
reconstruct $s$ from $t$, $b$, and $l$.  For example, given
the \IdentityLens{} lens, we can recover $s$ from $t$ using no bits
because in this case, $s$ is just $t$.  In contrast, given
the \Disconnect{} lens, we need enough bits to fully encode $s$ in
order to recover it from $t$ because $t$ contains no useful
information.  This calculation shows that if both \IdentityLens{}
and \Disconnect{} are satisfying lenses, then \IdentityLens{} is more
likely.

The expected number of bits required to recover data is a
well-known concept in information theory: it is the \emph{entropy} of
the data. Calculating the 
entropy of data requires a probability distribution over the
space of possible values for the data.  Specifically,
given a set $S$ and a probability distribution over that set $P
\OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
  S}P(s)\Log_2P(s)$.

@@

This infrastructure gives us what we need to find likely
satisfying lenses from a collection of satisfying lenses.  All such
lenses must synchronize data between the same two 
stochastic regular expressions $\Regex$ and $\RegexAlt$.
For any such lens $\Lens$, we can calculate the expected number of
bits required 
to recover a string $s$ in $\LanguageOf{\Regex}$ from a synchronized
string $t$ in $\LanguageOf{\RegexAlt}$. 
This expected number of bits is just the conditional entropy of
$\Regex$ given $\RegexAlt$ and $\Lens$. 
To ensure symmetry, the likelihood we assign to $\Lens$ is the sum of 
the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and
the conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$.

@@

Now that we have a way to calculate likelihood, we need a way to search for
it.
Previous work on synthesis~\ksf{add citations} has shown
that we can use types to dramatically restrict the search space and
improve the effectiveness of synthesis.  In our setting, types are
pairs of stochastic regular expressions, which come with a rich
algebraic structure that makes each regular expression semantically
equal to infinitely many others.  To tame that search space and
leverage the benefits of type-directed synthesis, we split our
algorithm into two communicating synthesizers following earlier
work~\cite{quotient-lens-synthesis}.  The first
synthesizer, \RXSearch, navigates the space of semantically equivalent
regular expressions by iteratively applying rewrite rules that
preserve both the semantics and the probability
distributions of those stochastic regular expressions. \RXSearch
ranks pairs of stochastic regular expressions 
by the number of rewrite rule applications required to obtain the pair
from the stochastic regular expressions given as input. It passes the
pairs off to the second synthesizer, \GreedySynth, in rank order with
the smallest first.

At a high level, \GreedySynth looks for lenses between a given pair of
stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the
highest likelihood by performing a type-directed search. 
In more detail, \GreedySynth first uses type equalities based on semiring 
equivalences (ignoring the star semiring equivalences that are managed
by \RXSearch) to normalize the
stochastic regular expressions provided by \RXSearch into 
\emph{stochastic n-ary DNF regular expressions}.  
Then, \GreedySynth uses the syntax of these n-ary DNF regular
expressions to find highly normalized lenses, in a
form we call \emph{simple symmetric n-ary DNF lenses}.  These lenses include
neither a composition operator nor a type equivalence rule.  These
restrictions mean that each simple symmetric n-ary DNF lens is typed
by a single pair of stochastic n-ary DNF regular expressions and 
\GreedySynth's search space for a given pair of regular expressions is
finite.  The variations that \GreedySynth 
considers include choosing between \IdentityLens{} and  \Disconnect
lenses and deciding how to match multiple occurrences of the same kind
of data, \EG, the two occurrences of \lstinline{name} in the
management and HR data formats (we want to map first names to first
names and last names to last names). 
Finally, \GreedySynth converts this lens 
back into a simple symmetric lens by converting the n-ary forms into
the binary forms provided in the surface language.

As an example of the partnership between \RXSearch and \GreedySynth,
consider the lens shown in Figure~\ref{fig:example_lens}. 
Inferring this lens requires \RXSearch to use a star semiring axiom to
rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
\lstinline{emp_salaries} into 
\lstinline{"" | (("\n" . emp_salary). ("\n" . emp_salary)$^*$)} 
so that \GreedySynth can find the appropriate lens.

Our strategy of communicating synthesizers gives us a way to
enumerate pairs of stochastic regular expressions of increasing
rank and to efficiently search through them, but it poses a
problem: when do we stop? We may have found a promising lens
between a pair of stochastic regular expressions, but a different 
pair we haven't yet discovered may give rise to an even better lens. 
Or, we may have found a promising lens between a given pair of
stochastic regular expressions, but the discovered lens is not 
acceptable to the users.

Our algorithm must resolve a three-way tension between the quality of
the inferred lens, the number of examples the user must provide to
eliminate unwanted lenses, and the amount of time it takes to return a
result.  For example, if the algorithm has already found the lens in
Figure~\ref{fig:example_lens}, we don't want to spend a lot of time
searching for an even better lens.

To resolve this tension, \RXSearch uses heuristics to judge whether to
return the current best satisfying lens to the user or to pass the
next set of unrollings to \GreedySynth.
The heuristics favor stopping if the current best satisfying
lens has relatively high likelihood, indicating the lens is very promising.
The heuristics also favor stopping if \RXSearch 
has delievered to \GreedySynth all pairs of stochastic regular
expressions at a given rank and there is a large number of pairs at the next
rank because searching through all such pairs will take a long time.
With this approach, the algorithm can quickly respond with a satisfying
lens with relatively high likelihood.  If users are unhappy with the result, they can refine the
search by supplying additional examples, which serve both to rule out
the previously proposed lens(es) and to reduce the size of the search
space by cutting down on the number of satisfying lenses.
